---
title: "Getting Fact View quaters into Tableau"
output: github_document
---

```{r setup}
pacman::p_load(tidyverse, lubridate)

```

Working with quarters in Tableau can be challenging with the way the ICPI Fact View datasets are structured. We'll walk through restructuring the dataset to allow you to work with it more easily in Tableau


## CREATE TEST DATA FRAME TO WORK WITH

The Fact View dataset starts out in a semi wide form initially. Let's import it and subset it slightly to make it more manageable to work with.

| psnu | indicator | fy2017q1 | fy2017q2 | fy2017q3 |
|------|-----------|----------|----------|----------|
| x    | y         | #        | #        | #        |

```{r, warning=FALSE, results='hide'}
  #import data (change this to wherever your OU x IM dataset is located)
    df_factview <- read_tsv("~/ICPI/Data/ICPI_FactView_OU_IM_20171115_v1_1.txt") 
```

```{r, warning=FALSE}
df_factview <- df_factview %>% 
  #rename all columns to lower case [persnal preference]
    rename_all(tolower) %>% 
  
  #filter to smaller subset to work with
    filter(indicator == "TX_NEW", disaggregate == "Total Numerator", operatingunit == "Rwanda") %>% 
  
  #limit variable columns we're working with 
    select(operatingunit, mechanismid, indicator, contains("fy2")) 

  #view dataset
    head(df_factview)

```

## CLEAN UP

To Tableau needs data in a long format (below) rather than wide (above). So, this means we'll need to melt it, or reshape it to a long format. Beore reshaping, we need to clean it up a bit first.


```{r}

df_clean <- df_factview %>%
  
  #remove APR and TARGETS from the dataset since they are not quarterly data
    select(-contains("apr"), -contains("targets")) %>%
  
  #remove the fy from the start of the value column names to make it easier to work with when we convert from string to a date
    rename_at(vars(contains("fy2")), funs(str_replace(.,"fy", "") ))
  
  #view dataset
    head(df_clean)
```


## MELT

Now that the data is cleaned up a bit, let's reshape it.

| psnu | indicator | period   | value |
|------|-----------|----------|-------|
| x    | y         | fy2017q1 | #     |
| x    | y         | fy2017q2 | #     |
| x    | y         | fy2017q3 | #     |
| x    | y         | fy2017q4 | #     |

```{r}
 
df_long <- df_clean %>% 
  
  #melt dataset
    gather(pd, val, starts_with("2"), na.rm = TRUE)

  #view dataset
    head(df_long)

```

## CONVERT QUATER TO DATE 

The other issue is that PEPFAR deals in fiscal quarter not calendar quarters. So, FY2017Q1 is starts Oct 2016, not Jan 2017. It's easier to work with dates. 

| psnu | indicator | period   | date       | value |
|------|-----------|----------|------------|-------|
| x    | y         | fy2017q1 | 2016-10-01 | #     |
| x    | y         | fy2017q2 | 2017-01-01 | #     |
| x    | y         | fy2017q3 | 2017-04-01 | #     |

```{r}
df_long_date <- df_long %>% 
  
  #add date
    mutate(date = yq(pd)  %m-% months(3)) %>% #to change from fiscal quarter to date, need to subtract 3 mo
  
  #reorder
    select(operatingunit, mechanismid, indicator, pd, date, val)

head(df_long_date)
```